QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-11-21 Revision/Paper_Rev.tex
Rev_V9.0

For energy-constrained MEC systems, reducing energy consumption becomes a key objective, motivating the creation of energy-efficient approaches. 
To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to optimize resource allocation and computation offloading concurrently. 
%%In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
To ensure that minimizing energy consumption does not lead to a reduction in the system's computing rate, some researches have expanded their work by incorporating the maximum tolerable delay for task models.
In \cite{chouikhi2023energy}, Chouikhi \textit{et al.} proposed a computation offloading approach based on DRL to minimize long-term energy consumption and maximize the number of tasks completed before their tolerant deadlines.
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network. 
Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
However, energy-efficient approaches~\cite{munir2021multi}--\cite{huang2021deadline}, even when considering deadline constraints~\cite{wu2024combining}--\cite{huang2021deadline}, do not adequately address the complexities of time-sensitive applications, where reducing delay can have significant effects on the user's QoE. Particularly, saving transmission energy consumption may reduce the transmission power of MDs and lead to poor performance for real-time applications.

To address time-sensitive applications in energy-constrained MEC systems, solving the relationship between delay and energy consumption is the primary optimization objective.
Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments. 
 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 Liu \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. 
 Although delay-energy trade-off approaches effectively address time-sensitive tasks, they are not suitable for applications with strict requirements on delay or energy consumption. This is challenging to address, as processing deadlines impact the load dynamics at the ENs, which in turn affect the delay of offloaded tasks.

Incorporating task processing deadlines for time-sensitive applications in energy-constrained MEC systems requires achieving an effective balance among delay, energy consumption, and computation rate. 
  In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.
 To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent DQN-based algorithm. 
  In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
   Some reasearch takes deadline-constraint tasks to ensure that the minimized trade-off be within the maximum tolerated delay.
  Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
  Wu \textit{et al.} in \cite{wu2023computation} introduced a stochastic game-based resource allocation in the SDN-based MEC network. They used an MDP and proposed a multi-agent RL method to minimize both energy consumption and processing delay. 
Nevertheless, QoE is a time-varying performance measure that reflects user satisfaction, where each user may have unique QoE requirements. According to the MDs' individual preferences, each delay, energy consumption, and computation rate may take precedence over two others. While studies \cite{chen2021drl}-\cite{gao2022large} have explored these factors, they overlook user-specific demands and fail to meet QoE requirements effectively.

  Different from aformentioned works []-[], we are motivated to effectively address QoE optimization, where QoE accounts for task completion rate, processing delay, and energy consumption. To effectively address personalized QoE optimization, we designed an adaptive balance for these factors according to the user's preferences, such as prioritizing reduced task delay or energy saving, which may vary over time. Each user may have unique QoE requirements at any given time.

>>>>

For energy-constrained MEC systems, reducing energy consumption becomes a key objective, motivating the creation of energy-efficient approaches. 
Munir \textit{et al.} \cite{munir2021multi}  investigated an energy dispatch problem of a self-powered wireless network with MEC capabilities and developed a semi-distributed approach using a multi-agent RL framework to minimize energy consumption.
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to optimize resource allocation and computation offloading concurrently. 
%%In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
To ensure that minimizing energy consumption does not lead to a reduction in the system's computing rate, researches have expanded their work by incorporating the maximum tolerable delay for task models.
In \cite{chouikhi2023energy}, Chouikhi \textit{et al.} proposed a computation offloading approach based on DRL to minimize long-term energy consumption and maximize the number of tasks completed before their tolerant deadlines.
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network. 
Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
However, energy-efficient approaches~\cite{munir2021multi}--\cite{huang2021deadline}, even when considering deadline constraints, do not adequately address the complexities of time-sensitive applications, where reducing delay can have significant effects on the user's QoE. Particularly, saving transmission energy consumption may reduce the transmission power of MDs and lead to poor performance for real-time applications.

To address time-sensitive applications in energy-constrained MEC systems, solving the relationship between delay and energy consumption is the primary optimization objective.
Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments. 
 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 Liu \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. 
 She \textit{et al.} in \cite{she2024efficient} investigated the computation offloading optimization problem and proposed DRL-based approach to minimize both the transmission delay and energy, based on POMDP.
  Although delay-energy trade-off approaches~\cite{liao2023online}--\cite{she2024efficient} effectively address time-sensitive tasks, they are not suitable for applications with strict requirements on delay or energy consumption.

Incorporating task processing deadlines for time-sensitive applications in energy-constrained MEC systems requires achieving an effective balance among delay, energy consumption, and computation rate. 
  In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.
 To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent DQN-based algorithm. 
  In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
  Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
  Wu \textit{et al.} in \cite{wu2023computation} introduced a stochastic game-based resource allocation in the SDN-based MEC network. They used an MDP and proposed a multi-agent RL method to minimize both energy consumption and processing delay. While studies \cite{gao2022large}--\cite{wu2023computation} have explored delay-sensitive applications with prossecing deadline and addressed energy-constrained MEC, they overlook user-specific demands and fail to meet QoE requirements effectively.

In contrast to aformentationed related works \cite{zhang2023offline}--\cite{wu2023computation}, which overlook key performance metrics, we are motivated to effectively address QoE optimization, where QoE accounts for task completion rate, processing delay, and energy consumption. Different from \cite{zhang2023offline}--\cite{li2022integrated}, \cite{munir2021multi}--\cite{dai2020edge}, and \cite{liao2023online}--\cite{she2024efficient}, which considered delay-tolerant tasks, we take into account delay-sensitive tasks with processing deadlines. This is challenging to address, as processing deadlines impact the load dynamics at the ENs, which in turn affect the delay of offloaded tasks. 
  To effectively address personalized QoE optimization, we designed an adaptive balance for these factors according to the user's preferences, such as prioritizing reduced task delay or energy saving, which may vary over time. Each user may have unique QoE requirements at any given time.

QoE is a time-varying performance measure that reflects user satisfaction, where each user may have unique QoE requirements. According to the MDs' individual preferences, each delay, energy consumption, and computation rate may take precedence over two others.
